tl20-05_berghei_pir_GeneSetEnrichmentQusage

### 2021-09-14 chapter3_results_bergheiPir/Manuscript_Github_Pberghei_CalculateTPM.Rmd
Split PCA/hclust/topGenes file
berg: stats comparison mero-asex-malegam heatmap and save proportional bar chart

@@

>>>>

Need to look more closely as to what I should and shouldn't remove from the hclust. E03_Asex.Schiz_3 looks like a well clustering sample, some of the 54 and 60h Liver samples do not look very similar to one another.

@@

## Saving the counts with new names/samples removed

```{r save_new_counts}
berghei_counts[,str_subset(colnames(berghei_counts), 'E06_Liv.6h')] <- NULL

col_stage <- colnames(berghei_counts)

colnames(berghei_counts) <- col_stage

berghei_counts$`E08_Liv.pre4h_1` <- NULL
berghei_counts$`E03_Asex.Schiz_1` <- NULL
berghei_counts$`E03_Asex.Schiz_3` <- NULL

write.csv(berghei_counts,
          'berghei_counts_newNames.csv')
```

### 2021-09-14 chapter3_results_bergheiPir/Manuscript_berghei_pir_GeneSetEnrichmentQusage.Rmd
berg: stats comparison mero-asex-malegam heatmap and save proportional bar chart

```{r read_count_data}

berg_tpm <- read_xlsx('Manuscript_Frontiers_SuppInfo_berghei.xlsx',
                      sheet = 'berg_TPM_allExpts_bioReps') %>% 
  dplyr::rename_with(.fn = ~ str_replace(.x, 
                                         pattern = 'Expt', 
                                         replacement = 'E'))

@@

berg_tpm <- column_to_rownames(berg_tpm, var = 'Geneid') %>% 
  select(!matches('BldMl|Gam_'))

@@

@@

```{r}
livmero_livpost48.results <- lapply(
  list(pir_geneset,
       ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Liv.Meroz-Liv.post48h',
                                gene_list,
                                n.points = 2^18)

@@

```{r}
maleGam_v_Asex.Mix.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Gam.Male-Asex.Mixed',
                                gene_list,
                                n.points = 2^18)

@@

```{r}
maleGam_v_Liv.Meroz.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Gam.Male-Liv.Meroz',
                                gene_list,
                                n.points = 2^18)

@@

```{r}
Asex.Mix_v_Liv.Meroz.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Asex.Mixed-Liv.Meroz',
                                gene_list,
                                n.points = 2^18)

@@

```{r}
berghei_comparison_stats_df <- bind_rows(
  lapply(c(maleGam_v_Asex.Mix.results, 
           maleGam_v_Liv.Meroz.results, 
           Asex.Mix_v_Liv.Meroz.results,
           maleVfemale, 
           post48livvearlyliv, 
           livmero_livpost48.results),
         getElement,
         'qstable')
) %>% 
  select(contrast, everything()) %>% 
  mutate(FDR = p.adjust(p.Value, method = 'fdr')) %>% 
  filter(FDR < 0.05) %>% 
  arrange(contrast)

max_logfold <- ceiling(max(max(berghei_comparison_stats_df$log.fold.change), 
                           abs(min(berghei_comparison_stats_df$log.fold.change))))

(
x <- ggplot(berghei_comparison_stats_df, 
            aes(x = contrast, 
                y = reorder(pathway.name, dplyr::desc(pathway.name)))) +
  xlab('Comparison') + ylab('Geneset') + 
  # scale_x_discrete(labels= c(1,2,3,4,6)) +
  geom_count(aes(
                 colour = log.fold.change), 
             na.rm = TRUE) +
  labs(colour = 'log2 Fold Change') +
  scale_color_gradientn(
    colours = rev(brewer.pal(11, "RdBu")),
    na.value="transparent",
    limits = c(-max_logfold,max_logfold),
    # limits = c(-0.5,1.25)
  ) +
  scale_size_area(max_size = 6) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = 'white'),
        plot.background = element_blank(),
        legend.background = element_rect(fill = 'transparent'),
        axis.text.x = element_text(angle = 90))
  )
```

>>>>

```{r read_count_data}
berg_tpm <- read.csv('berghei_counts_newNames.csv',
                     row.names = 1)

@@

berg_tpm <-  select(berg_tpm, 
                    !matches('BldMl|Gam_')) %>% 
  column_to_rownames(var = 'Geneid')

@@

pir_info <- read.table('../standard_files/Pberghei_pir_clades.txt',
                    header = TRUE)

@@

lapply(post48livvearlyliv, function(x) `$`(x,qstable))
```

```{r}
livmero_livpost48.results <- lapply(
  list(pir_geneset,
       ShortLong_list,
       subfam_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Liv.Meroz-Liv.post48h',
                                gene_list,
                                n.points = 2^18)

    list(results = qs.results.subfam, 
         qstable = qsTable(qs.results.subfam) %>% 
           mutate(contrast = qs.results.subfam$contrast))
  }
)
lapply(livmero_livpost48.results, function(x) `$`(x,qstable))
```

@@

```{r maleGam_v_Asex.Mix}
maleGam_v_Asex.Mix.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Gam.Male-Asex.Mixed',
                                gene_list,
                                n.points = 2^18)

@@

```{r maleGam_v_Liv.Meroz}
maleGam_v_Liv.Meroz.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Gam.Male-Liv.Meroz',
                                gene_list,
                                n.points = 2^18)

@@

```{r Asex.Mix_v_Liv.Meroz}
Asex.Mix_v_Liv.Meroz.results <- lapply(
  list(pir_geneset,
    subfam_list, 
    ShortLong_list), 
  function(gene_list){
    qs.results.subfam <- qusage(berg_tpm_log2plus1.mat,
                                lifeCycle_labels,
                                'Asex.Mixed-Liv.Meroz',
                                gene_list,
                                n.points = 2^18)

@@

```{r fdr}
berghei_comparison_stats_df <- bind_rows(
  lapply(c(maleGam_v_Asex.Mix.results, 
           maleGam_v_Liv.Meroz.results, 
           Asex.Mix_v_Liv.Meroz.results,
           maleVfemale, 
           post48livvearlyliv, 
           livmero_livpost48.results),
         getElement,
         'qstable')
) %>% 
  select(contrast, everything()) %>% 
  mutate(FDR = p.adjust(p.Value, method = 'fdr')) %>% 
  arrange(contrast)
```

```{r}
max_logfold <- ceiling(max(max(berghei_comparison_stats_df$log.fold.change), 
                           abs(min(berghei_comparison_stats_df$log.fold.change))))

(
  x <- ggplot(filter(berghei_comparison_stats_df,
                     FDR < 0.05), 
              aes(x = contrast, 
                  y = reorder(pathway.name, dplyr::desc(pathway.name)))) +
    xlab('Comparison') + ylab('Geneset') + 
    # scale_x_discrete(labels= c(1,2,3,4,6)) +
    geom_count(aes(
      colour = log.fold.change), 
      na.rm = TRUE) +
    labs(colour = 'log2 Fold Change') +
    scale_color_gradientn(
      colours = rev(brewer.pal(11, "RdBu")),
      na.value="transparent",
      limits = c(-max_logfold,max_logfold),
      # limits = c(-0.5,1.25)
    ) +
    scale_size_area(max_size = 6) +
    theme_bw() +
    theme(panel.grid.minor = element_blank(),
          panel.background = element_rect(fill = 'white'),
          plot.background = element_blank(),
          legend.background = element_rect(fill = 'transparent'),
          axis.text.x = element_text(angle = 90))
)
```

```{r mero_asex_malegam}
mero_asex_malegam_stats_df <- filter(berghei_comparison_stats_df,
                                     contrast %in% c('Asex.Mixed-Liv.Meroz',
                                                     'Gam.Male-Asex.Mixed',
                                                     'Gam.Male-Liv.Meroz'))

order_path.names <- ordering(unique(mero_asex_malegam_stats_df$pathway.name), 
                             order_vector = c('pir', 
                                              'L', 'L[[:digit:]]', 
                                              'S', 'S[[:digit:]]'))

mero_asex_malegam_stats_df <- mero_asex_malegam_stats_df[order(match(mero_asex_malegam_stats_df$pathway.name,order_path.names)),]

max_logfold <- ceiling(max(max(mero_asex_malegam_stats_df$log.fold.change), 
                           abs(min(mero_asex_malegam_stats_df$log.fold.change))))

mero_asex_malegam_stats_df_cast <- dcast(mero_asex_malegam_stats_df,
      contrast ~ pathway.name,
      value.var = 'log.fold.change') %>% 
  column_to_rownames(var = 'contrast')

mero_asex_malegam_stats_df_cast <- mero_asex_malegam_stats_df_cast[,order_path.names]

Heatmap(mero_asex_malegam_stats_df_cast,
        name = 'logFoldChange',
        cluster_rows = FALSE,
        cluster_columns = FALSE,
        rect_gp = gpar(col = "white", lwd = 2),
        cell_fun = function(j, i, x, y, width, height, fill){
          fdr_val <- mero_asex_malegam_stats_df[
            mero_asex_malegam_stats_df$contrast == rownames(mero_asex_malegam_stats_df_cast[i,]),
          ]$FDR[j]
          if(fdr_val<= 0.05){
            grid.text(
              sprintf("%.3f", #One significant figure
                      fdr_val),
              x,
              y,
              gp = gpar(fontsize = 10)
            )
          }
        })
